\protect Connecting every bit of knowledge: \\ The structure of Wikipedia's First Link Network

### 2016-11-21 writeup/2016-04wikipedia-network/wikipedia-network.abs.tex
added pesudo-code and fixed centrality table

Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in each article, we algorithmically construct a directed network of 
all 4.7 million articles: Wikipedia's First Link Network.
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
articles on inventions, places, people, objects, and events are related and organized.  
By traversing every path, we measure the accumulation of first links, path lengths, basins, cycles, 
and the influence each article exerts in shaping the network.
We find scale-free distributions describe path length, 
accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
at a few articles---flowing from specific to general and culminating around fundamental notions such as
Community, State, and Science. 
Philosophy directs more paths than any other article by two orders of magnitude. 
We also observe a gravitation towards topical articles such as 
Health Care and Fossil Fuel. 
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

>>>>

Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in each article, we algorithmically construct a directed network of 
all 4.7 million articles: Wikipedia's First Link Network.
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
articles on inventions, places, people, objects, and events are related and organized.

By traversing every path, we measure the accumulation of first links, path lengths,
groups of path-connected articles, and cycles.
We also develop a new method, traversal funnels, to measure the influence each article exerts in shaping the network. 
Traversal funnels provides a new measure of influence for directed networks without spill-over into cycles, in contrast to traditional network centrality measures.
Within Wikipedia's First Link Network, we find scale-free distributions describe path length, 
accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
at a few articles---flowing from specific to general and culminating around fundamental notions such as
Community, State, and Science. 
Philosophy directs more paths than any other article by two orders of magnitude. 
We also observe a gravitation towards topical articles such as 
Health Care and Fossil Fuel. 
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

### 2016-11-21 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
added pesudo-code and fixed centrality table

Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
Informed by river network metrics,
we consider the FLN as a flow, quantifying 
the accumulation of first links around articles 
and the influence an article exerts in shaping the FLN.
Together with cycles, in-degree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

\begin{figure*}[tp!]
  \includegraphics[width=\textwidth]{traversal_visit_algo_figure.pdf}  
  \caption{
    \textbf{Traversal Visit Algorithm on a sample network.}
     The traversal visit vectors are an adjacency matrix for the paths through the network: 
     the first column indicates the path formed starting with article A. 
     The number of traversal visits for article A is then the number of paths containing A 
     or the sum of the first row in our matrix:
     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
  }
  \label{fig:Traversal Visits}
\end{figure*}
Starting at each article, we construct a path through the FLN, 
to map the flow of connections among articles.
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Previous studies have used flow to characterize the structure of river networks
\cite{horton1945erosional, dodds1999unified} and describe the organization of 
blood networks, food systems, and transportation networks
\cite{garlaschelli2003universal}.

@@

To compute traversal funnels on a directed network, begin at each node: 1. Traverse the edge to the next node 2. Increment the next node's count by one 3. Repeat while the next node is unvisited. The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle.

@@

Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
In addition, the work provides a general algorithm for assessing the structure and measuring influence in a directed network (or any braided flow structure).
These findings also form the basis for the creation of a taxonomy where 
every idea, event, or object sits within a hierarchy of connected notions.
The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
Specifically, an ever-evolving reference of related hierarchical concepts could be 
used to improve search engine algorithms or natural language processing.

>>>>

Our goal is to study the structure of the FLN for insight into how the information on Wikipedia is organized and related.
Informed by river network metrics,
we consider the FLN as a flow. We quantify
the accumulation of first links around articles 
and develop a new method, \textit{traversal funnels}, to measure the influence an article exerts in shaping the FLN.
Together with cycles, in-degree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

\begin{figure*}[tp!]
  \includegraphics[width=\textwidth]{traversal_visit_algo_figure.pdf}  
  \caption{
    \textbf{Traversal Visit Algorithm on a sample network.}
     The traversal visit vectors are an adjacency matrix for the paths through the network: 
     the first column indicates the path formed starting with article A. 
     The number of traversal visits for article A is then the number of paths containing A 
     or the sum of the first row in our matrix:
     $\sum_{i=1}^7 A_{\text{visit, i}} = 7$.
  }
  \label{fig:Traversal Visits}
\end{figure*}
Starting at each article, we construct a path through the FLN, 
to map the flow of connections among articles.
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Previous studies have used flow to characterize the structure of river networks
\cite{horton1945erosional, dodds1999unified} and describe the organization of 
blood networks, food systems, and transportation networks
\cite{garlaschelli2003universal}. 
In the same vein, we use flow to measure accumulation and develop a new method for isolating influence in a directed network with cycles.

@@

The algorithm increments the traversal funnels count along each path, up to the start of a cycle. The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle. Below is the pseudo-code.

\begin{verbatim}
Traversal Funnels Algorithm

// 1. Determine Cycles 
for node in Graph
    next_node = node.link
    visited = {node, next_node}
    // advance to potential cycle
    while next_node not in visited:
        next_node = next_node.link
        visited.add(next_node)
    // mark cycle
    if (node == next_node):
        node.in_cycle = True

// 2. Compute Traversal Funnels
for node in Graph:
    while node.in_cycle == False:
        node.funnels += 1
        node = node.link
\end{verbatim}

The algorithm's average runtime is proportional to the product of the number of nodes, $n$, and the average path length, $l$. If the average path length is fixed (independent of $n$), the runtime is $O(n)$ with $O(n + l)$ auxiliary memory for storing cycles and traversing paths. In the worst case where $l$ is proportional to $n$, the runtime is $O(n^2)$ with $O(n)$ auxiliary memory for storing cycles and traversing paths. In implementation, each iteration step in the algorithm can be parallelized since the algorithm is order agnostic, as long as each node is eventually selected.

@@

Future work could examine other language versions of Wikipedia for potentially telling cultural or regional differences as well as expand the network to more than the first link.
In addition, the work provides a general algorithm for assessing the structure and measuring influence in a directed network (or any braided flow structure). 
Traversal funnels is a new measure of influence for directed networks without spill-over into cycles in contrast to traditional network centrality measures.
When applied to Wikipedia, these findings also form the basis for the creation of a taxonomy where 
every idea, event, or object sits within a hierarchy of connected notions.
The taxonomy would extend a traditional word thesaurus beyond mere synonyms to a related hierarchy of concepts.
Applications could range from an enhanced network of ideas to psychological insights into how humans form associations.
Specifically, an ever-evolving reference of related hierarchical concepts could be 
used to improve search engine algorithms or natural language processing.
